---
title: "Architecture and compatibility contracts"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Architecture and compatibility contracts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(FielDHub)
```

## Layers and direction of dependencies

| Layer | Responsibilities | Must not do |
|---|---|---|
| Public engines (`fct_`, `engine_`, `validate_`) | Validate inputs, randomize, optimize, construct results | Depend on Shiny, display dialogs, install packages, change caller options |
| Results (`result_`) | Schema, declarative engine registry, replay, S3 presentation | Evaluate function names or handlers supplied by saved metadata |
| Layout/render/simulation (`layout_`, `render_`, `sim_`) | Coordinates, drawing, numerical simulation | Re-randomize a saved design merely to display it |
| I/O (`io_`) | Parse uploads, validated format dispatch, workflow archives | Treat a CSV field book as a complete replay record |
| App (`app_`) | Declarative page controls, argument builders, one generic module | Duplicate engine logic or mutate the accepted result when controls change |

Script users need only Imports. `run_app()` checks its optional packages and
reports an installation command if needed; it does not install anything.
`source("dev/run_dev.R")` is the source-development launcher, while deployed
images start the installed package. Neither requires a golem config file.

Every app request is converted to a plain named argument list and sent to a
whitelisted public engine. A completed design is an immutable snapshot for
layout, simulation and downloads. With opt-in `workers > 0`, installed-package
jobs use a private mirai pool; source development remains synchronous. Workers
close when the app stops. The 100 MiB process-wide Shiny upload setting is scoped
to the app lifetime and restored afterward, not on each session disconnect.

## Results and randomness

A design has class `c("fieldhub_<key>", "FielDHub")`, `infoDesign`, `fieldBook`
and schema-1 metadata. The common book keys are `ID`, `LOCATION`, and `PLOT`;
the registry declares additional columns for each design. Extra columns are
allowed. Field coordinates belong to the layout contract; some engines record
them at construction, while others offer multiple layouts. `PLOT` alone is not
a universal unique key for split units, fillers or multiple locations.

Metadata records the design key, schema version, effective seed, RNG kinds,
package version and effective parameters. An explicit seed restores the caller
stream; an omitted seed consumes exactly one automatically selected seed, then
isolates all internal randomization. Failed input validation should occur before
that draw. Numerical optimizers can also depend on R, dependencies and platform;
metadata is not a guarantee of cross-version bit-identical reconstruction.

```{r recorded-result}
x <- latin_rectangle(5, rows = 3, seed = 27)
x$metadata[c("design", "schema_version", "seed", "package_version")]
stopifnot(identical(reproduce_design(x), x))
```

Allocation plans and pair-swap results have separate validated classes but use
the same metadata/replay boundary. Legacy 1.5.x results still have their
documented printing and plotting path. Missing historical parameters cannot be
recovered automatically and must not be invented.

## Extension boundaries

The reviewed `fieldhub_design_registry()` is the single declaration of engine
names and field-book extensions. New fixed-coordinate designs can declare
generic layout/render handlers and a title, avoiding new dispatcher branches.
Existing specialized S3 methods preserve established layouts and output.
Optimization strategies live in plain engine helpers and record their choices
as data; visualizations use the layout/render S3 boundary. Neither should leak
UI concerns into scientific computation.

These are contributor interfaces for package code, not a mutable global plugin
registry. External packages should use the documented public engines,
`field_layout()`, replay and I/O APIs rather than depend on internal function
names. Adding runtime third-party handlers would need a separate compatibility
and security proposal. See the *Adding designs and formats* vignette for the
concrete extension procedure and the Latin rectangle example.

## Versioned interchange

```{r formats}
design_formats()
file <- tempfile(fileext = ".rds")
write_design(x, file)
stopifnot(identical(read_design(file), x))
unlink(file)
```

Format version 1 and result schema version 1 are independent contracts. RDS uses
R serialization version 2 and preserves the exact result. The validated reader
rejects unknown result schemas and engines and does not replay or execute code.
Only deserialize files from trusted sources. An R script is a write-only replay
artifact, never an importer; inspect it before explicitly sourcing it.
Script export declines oversized uploads/nonportable parameters rather than
emitting a misleading incomplete program. App workflow ZIPs separately carry
layout selections, simulated values, CSV data and metadata.

No existing file is overwritten by `write_design()` unless `overwrite = TRUE`.
Serialization happens first in a temporary file; the final copy is not an
atomic filesystem transaction. Preserve backups of important outputs.

## Compatibility and release ownership

Scientific outputs for a fixed seed are part of the compatibility contract.
Correctness fixes may change them, but require a regression, an independent
scientific check, an explained snapshot diff and a NEWS entry naming affected
inputs. Do not silently update snapshots to make a refactor pass. New public
options retain old defaults unless a deliberate change is documented.

Deprecated public arguments retain an alias and a classed warning through at
least one published minor release before removal is proposed. A schema change
requires a version increment and an explicit migration/rejection policy.
Internal extension changes must update the contributor guide and contract tests
in the same change; they are not an undocumented third-party API promise.

Maintainers choose a release owner and an independent backup/reviewer for each
release. Green checks do not confer scientific or publication approval. The
repository's `RELEASING.md` requires reviewed benchmarks, a platform/R matrix,
deployment evidence, licensing resolution, NEWS and an exact release tag.
Local verification does not establish remote CI or CRAN acceptance.

Feedback is opt-in through issues or the published maintainer contact. Request
minimal synthetic reproductions; do not collect private field books or add
telemetry. Review/release roles and contributor identities require agreement,
not inference from activity counts or email similarity.
