---
title: "Full Factorial Design"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Full Factorial Design}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = NA,
  warning = FALSE, 
  message = FALSE
)
```

# App

## Inputs

1. **Import entries' list?** Choose whether to import a list with entry numbers and names for genotypes or treatments.
    * If the selection is `No`, that means the app is going to generate synthetic data for entries and names of the treatment/genotypes based on the user inputs.

    * If the selection is `Yes`, the entries list must fulfill a specific format and must be a `.csv` file. The file must have the single column `TREATMENT`, containing a list of unique names that identify each treatment/genotype. Duplicate values are not allowed, all entries must be unique. In the following, we show an example of the entries list format. This example has an entry list with four checks and 8 treatments/genotypes. It is crucial to allocate the checks in the top part of the file.

```{r, include=FALSE}
library(FielDHub)
TREATMENT <- c(paste0("TRT_", LETTERS[1:10]))
df <- data.frame(TREATMENT)
```

```{r, echo = FALSE, results='asis'}
library(knitr)
kable(df)
```

2. Choose whether to use the factorial design in a RCBD or CRD with the **Select a Factorial Design Type** box.

3. Set the number of entries for each factor in a comma separated list in the **Input # of Entries for Each Factor** box. We want our example experiment to have 3 factors with 2, 3, and 4 entries respectively, so enter 2,3,4 in the box. 

4. Set the number of replications of the squares with the **Input # of Full Reps** box. 

5. Enter the number of locations in **Input # of Locations**. 

6. Enter the starting plot number in the **Starting Plot Number** box. If the experiment has multiple locations, you must enter a comma separated list of numbers the length of the number of locations for the input to be valid. 

7. Optionally, you may enter a name for the location of the experiment in the **Input Location** box. 

8. Select `serpentine` or `cartesian` in the **Plot Order Layout**. For this example we will use the default `serpentine` layout.

9. As with all the designs, we can set a seed number in the box labeled **Seed Number**. In this example, we will set it to 123. 

10. Once we have entered the information for our experiment on the left side panel, click the **Run!** button to run the design. 

## Outputs
After you run a full factorial design in FielDHub, there are several ways to display the information contained in the field book. 

### Field Layout
When you first click the run button on a full factorial design, FielDHub displays the Field Layout tab, which shows the entries and their arrangement in the field. In the box below the display, you can change the layout of the field or change the location displayed. 
You can also display a heatmap over the field by changing **Type of Plot** to **Heatmap**. To view a heatmap, you must first simulate an experiment over the described field with the **Simulate!** button. A pop-up window will appear where you can enter what variable you want to simulate along with minimum and maximum values. 

### Field Book
The **Field Book** displays all the information on the experimental design in a table format. It contains the specific plot number and the row and column address of each entry, as well as the corresponding treatment on that plot. This table is searchable, and we can filter the data in relevant columns. If we have simulated data for a heatmap, an additional column for that variable appears in the Field Book. 

# Function : `full_factorial()`

You can run the same design with a function in the FielDHub package, `full_factorial()`. We can enter the information describing the above design like this:

```{r, echo = TRUE, warning=FALSE, comment=''}
factorial <- full_factorial(setfactors = c(2,3,4), #numeric vector with levels of each factor
                            reps = 3, #number of replicates
                            l = 1, #number of locations
                            type = 2, #CRD or RCBD, 1 or 2 respectively
                            plotNumber = 101,
                            planter = "serpentine",
                            locationNames = "FARGO",
                            seed = 123) #seed number
```
```{r, comment='', echo=FALSE}
print(factorial)
```
This returns a list consisting of all the information displayed in the output tabs in the FielDHub app: design information, plot layout, plot numbering, entries list, and field book. These are accessible by the `$` operator, i.e. `factorial$layoutRandom` or `factorial$fieldBook`. 

`factorial$fieldBook` is a list containing information about every plot in the field, with information about the location of the plot and the treatment in each plot. As seen in the output below, the field book has columns for `ID`, `LOCATION`, `PLOT`, `REP`, and `TRT_COMB`, and columns for each factor individually.
```{r, comment='', echo=FALSE}
print(head(factorial$fieldBook))
```