---
title: "Optimized Multi-Location P-rep Design"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Optimized Multi-Location P-rep Design}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
options(rmarkdown.html_vignette.check_title = FALSE)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = NA,
  warning = FALSE, 
  message = FALSE
)
```

```{r, echo = FALSE}
library(FielDHub)
library(magrittr)
library(knitr)
library(kableExtra)
```

This vignette shows how to generate an **optimized multi-location partially replicated design** 
using both the FielDHub Shiny App and the scripting function `multi_location_prep()` from 
the `FielDHub` R package.

## Overview

Partially replicated designs are frequently used in initial generation field trials. This design type involves replicating a portion of the entries, while the remaining entries only appear once in the experiment. The optimized partially replicated (p-rep) design utilizes a genotype allocation matrix to determine the distribution of replicated and non-replicated entries across multiple locations, employing the principles of incomplete blocks
designs (IBD).

## Use case

Suppose there is a plant breeding field trial with 150 entries to be tested across five environments, where up to seven replications of each entry are allowed. Additionally, the project includes three checks; each replicated six times. We can generate an optimized multi-location partially replicated design using these parameters. This strategy guaranty that all treatments are present in all environments but in different amounts of replications. We can generate this design using the FielDHub Shiny app and the FielDHub multi_location_prep() standalone function in R.

## Running the Shiny App

To launch the app you need to run either

```{r, eval=FALSE}
FielDHub::run_app()
```

or

```{r, eval=FALSE}
library(FielDHub)
run_app()
```

## 1. Using the FielDHub Shiny App

Once the app is running, click the tab **Partially Replicated Design** and 
select **Optimized Multi-Location p-rep** from the dropdown. 

Then, follow the following steps where we will show how to generate an optimized 
partially replicated design.

## Inputs

1. **Import entries' list?** Choose whether to import a list with entry numbers and names for genotypes 
or treatments.

    * If the selection is `No`, that means the app is going to generate synthetic data for entries 
    and names of the treatment/genotypes based on the user inputs.

    * If the selection is `Yes`, the entries list must fulfill a specific format and must be a `.csv` file. 
    The file must have the columns `ENTRY` and `NAME`. The `ENTRY` column must have a unique entry 
    integer number for each treatment/genotype. The column `NAME` must have a unique name that 
    identifies each treatment/genotype. Both `ENTRY` and `NAME` must be unique, duplicates 
    are not allowed. In the following table, we show an example of the entries list format. 
    
```{r, include=FALSE}
ENTRY <- 1:10
NAME <- c(paste0("Genotype", 1:10))
df <- data.frame(ENTRY,NAME)
```

```{r, echo = FALSE, results='asis'}
df %>%
  kbl() %>%
  kable_styling()
```

2. Enter the number of entries in the **Input # of Entries** box as a comma separated list. 
In our example we will have 150 entries, so we enter 150 in the box for our sample experiment. 

3. Select whether or not the experiment will contain checks under the **Include checks?** option. 
The example experiment does, so set this to `Yes`. 

4. Once we select `Yes` on the above option, two more boxes appear, the first being **Input # of Checks** 
where we set how many checks to include in the experiment. In our case this is 3. 

5. Next to this option we have **Input # Check's Reps**, where we set the number of replications for 
each check respectively in a comma separated list. We are replicating each of the 3 checks 6 times, 
so enter `6,6,6` in this box. 

6. Enter the number of locations in **Input # of Locations**. We will run this experiment over 5 locations, 
so set **Input # of Locations** to 5.

7. Set the total number of replications of the entries over all locations in the `# of Copies Per Entry` 
dropdown box. For this example experiment, set this to 7. 

8. Select `serpentine` or `cartesian` in the **Plot Order Layout**. For this example we will use the 
default `serpentine` layout.

9. To ensure that randomizations are consistent across sessions, we can set a random seed in the 
box labeled **Random Seed**. In this example, we will set it to `2456`. 

10. (Optional) Enter the starting plot number in the **Starting Plot Number** box. Since the experiment has multiple locations, you must enter a comma separated list of numbers the length of the number of locations for 
the input to be valid. In this example, we will set it as `1,1001,2001,3001,4001`.

11. (Optional) Enter the location names in the **Input Location Name** box. Since the experiment has six 
locations, you must enter a comma separated list of strings for the names of the environments. In this 
example, we will set it as `LOC1,LOC2,LOC3,LOC4,LOC5`.

12. Once we have entered the information for our experiment on the left side panel, click the **Run!** 
button to run the design. 

13. You will then be prompted to select the dimensions of the field from the list of options in the 
dropdown in the middle of the screen with the box labeled **Select dimensions of field**. 
In our case, we will select `12 x 19`. 

14. Click the **Randomize!** button to randomize the experiment with the set field dimensions and 
to see the output plots. If you change the dimensions again, you must re-randomize.

If you change any of the inputs on the left side panel after running an experiment initially, 
you have to click the Run and Randomize buttons again, to re-run with the new inputs.

## Outputs

After you run a `Optimized Multi-Location P-rep Design` in FielDHub and set the dimensions of the field, there
are several ways to display the information contained in the field book. The first tab, **Get Random**, 
shows the option to change the dimensions of the field and re-randomize, as well as the genotype 
allocation matrix generated for the optimized p-rep design, which displays the replications of each 
genotype over each location, much like the matrix generated in sparse allocation. 

### Randomized Field

The **Randomized Field** tab displays a graphical representation of the randomization of the 
entries in a field of the specified dimensions. The replicated entries are the green colored 
cells, with the which cells appearing only once in the location. The display includes numbered 
labels for the rows and columns. You can copy the field as a table or save it directly as an 
Excel file with the _Copy_ and _Excel_ buttons at the top. 

### Plot Number Field

On the **Plot Number Field** tab, there is a table display of the field with the plots numbered 
according to the Plot Order Layout specified, either _serpentine_ or _cartesian_. You can see 
the corresponding entries for each plot number in the field book. Like the Randomized Field tab, 
you can copy the table or save it as an Excel file with the _Copy_ and _Excel_ buttons. 

### Field Book

The **Field Book** displays all the information on the experimental design in a table format. 
It contains the specific plot number and the row and column address of each entry, as well as 
the corresponding treatment/genotype on that plot. This table is searchable, and we can filter the data 
in relevant columns. 

## 2. Using the `FielDHub` function: `multi_location_prep()`.

You can run the same design with the function `multi_location_prep()` in the `FielDHub` package.

First, you need to load the `FielDHub` package typing,

```{r, echo = TRUE}
library(FielDHub)
```
Then, you can enter the information describing the above design like this:

```{r, echo = TRUE}
optim_multi_prep <- multi_location_prep(
  lines = 150, 
  l = 5, 
  copies_per_entry = 7, 
  checks = 3,
  rep_checks = c(6,6,6),
  plotNumber = c(1,1001,2001,3001,4001),
  locationNames = c("LOC1", "LOC2", "LOC3", "LOC4", "LOC5"),
  seed = 2456
)
```

#### Details on the inputs entered in `optimized_arrangement()` above

The description for the inputs that we used to generate the design,

*   `lines = 150` is the number of entries in the field.
*   `l = 5` is the number of locations.
*   `copies_per_entry = 7` is the number of copies of each entry. 
*   `checks = 3` (Optional) is the number of checks.
*   `rep_checks = c(6,6,6)` (Optional) is the number of replications of each check, 
    in a vector the length of the number of checks. 
*   `locationNames = c("LOC1", "LOC2", "LOC3", "LOC4", "LOC5")` are optional names for the        locations.
*   `seed = 1234` is the random seed to replicate identical randomizations.

### Print `optim_multi_prep` object

To print a summary of the information that is in the object `optim_multi_prep`, we can use the generic 
function `print()`. 

The `sparse_allocation()` function returns a list of 4 objects, `designs`, `list_locs`, `allocation`, 
and `size_locations`. The object `designs` contains the randomization and information about the design 
parameters for each location, while `list_locs` is a list of data frames. Each data frame has two columns; `ENTRY` and `NAME` with the information to randomize to each environment. The object `allocation` is the binary allocation matrix of genotypes to locations, and `size_locations` is a data frame with a column for each location and a row indicating the size of the location (number of field plots). 

For example, we can display the `allocation` object. Let us print the first ten genotypes allocation.

```{r, echo=TRUE, eval=FALSE}
print(head(optim_multi_prep$allocation, 10))
```

```{r, echo=FALSE, eval=TRUE}
print(head(optim_multi_prep$allocation, 10))
```

Let us add two new columns to the allocation table. We can ad the number of copies by genotype,
it should be 7 for all of them. Also we can add the average allocation by genotype, we can see each 
treatment will appear 1.4 times in average.

```{r, echo=FALSE, eval=TRUE}
optim_multi_prep$allocation %>%
        dplyr::mutate(
            Copies = rowSums(.),
            Avg = Copies / 5
        ) %>%
        head(10) %>%
        `rownames<-`(paste0("Gen-", 1:10)) %>%
        kbl() %>%
        kable_styling()
```

We can manipulate the sparse_allocation object as any other list in R. For example, we can save the 
design randomizations to an object called `designs`, and then index based on location  using the dollar
sign operator. We can first display the design parameters for LOC1 with the following code:

```{r, echo=TRUE, eval=FALSE}
design_randomization <- optim_multi_prep$designs
print(design_randomization$LOC1)
```

which outputs:

```{r, echo=FALSE, eval=TRUE}
design_randomization <- optim_multi_prep$designs
print(design_randomization$LOC1)
```

And for environment 2 by changing the location to print to `LOC2`:

```{r, echo=TRUE, eval=FALSE}
print(design_randomization$LOC2)
```

which outputs:

```{r, echo=FALSE, eval=TRUE}
print(design_randomization$LOC2)
```

### Access to `optim_multi_prep` output

The object `designs` we saved `optim_multi_prep$designs` to is a list consisting of all the information 
displayed in the output tabs in the FielDHub app: design information, plot layout, plot numbering, 
entries list, and field book, indexed by each location in the experiment. These are accessible by 
the `$` operator, i.e. `designs$LOC1$layoutRandom` or `designs$LOC1$fieldBook` for LOC1, etc. 

`designs$LOC1$fieldBook` is a data frame containing information about every plot in the field, 
with information about the location of the plot and the treatment in each plot. As seen in the 
output below, the field book has columns for `ID`, `EXPT`, `LOCATION`, `YEAR`, `PLOT`, `ROW`, 
`COLUMN`, `CHECKS`, `ENTRY`, and `TREATMENT`.

Let us see the first 10 rows of the field book for the first location in this experiment.

```{r, echo=TRUE, eval=FALSE}
field_book <- design_randomization$LOC1$fieldBook
head(field_book, 10)
```

```{r, echo=FALSE, eval=TRUE}
field_book <- design_randomization$LOC1$fieldBook
head(field_book, 10)
```

### Plot field layout

For plotting the layout in function of the coordinates `ROW` and `COLUMN` in the field book object
we can use the generic function `plot()` as follows. This plots only the first location, but this 
is indexable by location using the dollar sign operator as well. 

#### Field Layout for Location 1

```{r, fig.align='center', fig.width=7.2, fig.height=5.5}
plot(design_randomization$LOC1)
```

In the figure above, green plots contain replicated entries, and gray plots contain entries 
that only appear once.

#### Field Layout for Location 5

Also, for example the location five:

```{r, fig.align='center', fig.width=7.2, fig.height=5.5}
plot(design_randomization$LOC5)
```

<br>
<br>
<br>

