---
title: "Optimized Arrangement"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{optimized}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  options(rmarkdown.html_vignette.check_title = FALSE)
---

# Optimized Arrangement in FielDHub App
One kind of unreplicated design you can use in FielDHub is the optimized arrangement. In this design you can set the number of entries and the number of checks for the experiment. You can also choose to run the same experiment over multiple locations. 
In this vignette, we will show an example of an optimized arrangement design experiment over 3 locations with 3 checks of 12, 10, and 8 reps each, and 405 entries. We will distribute the total 435 plots over a 29 x 15 field. FielDHub will return a and a table with all the field information for all locations, with an entry for every plot. 

First, go to **Unreplicated Designs** > **Optimized Arrangement**

##  Inputs 

1. **Import entries' list?** Choose whether to import a list with entry numbers and names for genotypes or treatments.
    * If the selection is `No`, that means the app is going to generate synthetic data for entries and names of the treatment/genotypes based on the user inputs.

    * If the selection is `Yes`, the entries list must fulfill a specific format and must be a `.csv` file. The file must have the columns `ENTRY` and `NAME`. The `ENTRY` column must have a unique entry integer number for each treatment/genotype. The column `NAME` must have a unique name that identifies each treatment/genotype. Both ENTRY and NAME must be unique, duplicated are not allowed. As follow, we show an example of the entries list format. This example has an entry list with four checks and 8 treatments/genotypes. It is crucial to allocate the checks in the top part of the file.

2. Enter the number of checks in the **Input # of Checks** box, which is 3 in our case.

3. Enter the number of replications of the checks in a comma separated list containing a number for each check in the **Input # Check's Reps** box. For our example experiment, we will enter 10,10,10.

4. Enter the number of entries/treatments in the **Input # of Entries** box, which is 300 in our case.

5. Select `serpentine` or `cartesian` in the **Plot Order Layout**. For this example we will set the `serpentine` layout.

6. Since we want to run this experiment over 3 locations, set **Input # of Locations** to 3.

7. Enter the name of the site/location in the **Input the Location** box. For this experiment we will set the site as `FARGO`. In the case of users will run the experiment in multiple locations, the name for each location must be enter separete by comma, for example: `FARGO, CASSELTON, MINOT`.

8. To ensure that randomizations are consistent across sessions, we can set a seed number in the box labeled **Seed Number**. For instance, we will set it to 123. 

9. Enter the starting plot number in the **Starting Plot Number** box. In this experiment we want the plot start at `1001`.

10. Once we have entered the information for our experiment on the left side panel, click the **Run!** button to run the design. 

11. You will then be prompted to select the dimensions of the field from the list of options in the drop down in the middle of the screen with the box labeled **Select dimensions of field**. In our case, we will select 29 x 15. 

12. Click the **Randomize!** button to randomize the experiment with the set field dimensions and to see the output plots. If you change the dimensions again, you must re-randomize.

If you change any of the inputs on the left side panel after running an experiment initially, you have to click the Run and Randomize buttons again, to re-run with the new inputs.

## Outputs
After you run a single diagonal arrangement in FielDHub and set the dimensions of the field, there are several ways to display the information contained in the field book. The first tab, **Get Random**, shows the option to change the dimensions of the field and re-randomize, as well as a reference guide for experiment design. 

### Data Input
On the second tab, **Data Input**, you can see all the entries in the randomization in a list, as well as a table of the checks with the number of times they appear in the field. In the list of entries, the reps for each check is included as well.


### Randomized Field
The **Randomized Field** tab displays a graphical representation of the randomization of the entries in a field of the specified dimensions. The checks are all colored uniquely, showing the number of times they are distributed throughout the field. The display includes numbered labels for the rows and columns. You can copy the field as a table or save it directly as an Excel file with the _Copy_ and _Excel_ buttons at the top. 

### Plot Number Field
On the **Plot Number Field** tab, there is a table display of the field with the plots numbered according to the Plot Order Layout specified, either _serpentine_ or _cartesian_. You can see the corresponding entries for each plot number in the field book. Like the Randomized Field tab, you can copy the table or save it as an Excel file with the _Copy_ and _Excel_ buttons. 

### Field Book
The **Field Book** displays all the information on the experimental design in a table format. It contains the specific plot number and the row and column address of each entry, as well as the corresponding treatment on that plot. This table is searchable, and we can filter the data in relevant columns. 

# Function : `optimized_arrangement()`

You can run the same design with a function in the FielDHub package, `optimized_arrangement()`. We can enter the information describing the above design like this:
```{r, echo = FALSE, warning=FALSE, comment='', include=FALSE}
library(knitr)
library(FielDHub)
optim_expt <- optimized_arrangement(
  nrows = 29, 
  ncols = 15, 
  lines = 405, 
  amountChecks = c(12,10,8), 
  checks = 3,
  l = 3,
  plotNumber = c(1001,2001,3001),
  seed = 123,
  exptName = "Expt1"
)

```

```{r, results='hide'}
optim_expt <- optimized_arrangement(
  nrows = 29, #number of rows in the field, 29 x 15
  ncols = 15, #number of columns in the field
  lines = 405, #number of entries
  amountChecks = c(12,10,8), #numeric vector representing 
                             #respective replicates of each check, 
                             #or an integer total number of checks
  checks = 3, #number of checks
  l = 3, #number of locations
  plotNumber = c(1001,2001,3001), #numeric vector containing 
                                  #the starting plot number 
                                  #for each location respectively, 
                                  #or a single number for 1 location
  seed = 123, #seed number to replicate identical randomizations
  exptName = "Expt1" #optional name for experiment
)
```


```{r, comment='',echo=FALSE}
print(optim_expt)
```
This returns a list consisting of all the information displayed in the output tabs in the FielDHub app: design information, plot layout, plot numbering, entries list, and field book. These are accessible by the `$` operator, i.e. `optim_expt$layoutRandom` or `optim_expt$fieldBook`. 

`optim_expt$fieldBook` is a list containing information about every plot in the field, with information about the location of the plot and the treatment in each plot. As seen in the output below, the field book has columns for `ID`, `EXPT`, `LOCATION`, `YEAR`, `PLOT`, `ROW`, `COLUMN`, `CHECKS`, `ENTRY`, and `TREATMENT`.
```{r, comment='', echo=FALSE}
print(head(optim_expt$fieldBook))
```

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
