---
title: "Augmented Randomized Complete Block Design"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{rcbd_a}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE, 
  message = FALSE
)
```

# Augmented RCBD in FielDHub App

## Inputs

1. **Import entries' list?** Choose whether to import a list with entry numbers and names for genotypes or treatments.
    * If the selection is `No`, that means the app is going to generate synthetic data for entries and names of the treatment/genotypes based on the user inputs.

    * If the selection is `Yes`, the entries list must fulfill a specific format and must be a `.csv` file. The file must have the columns `ENTRY`, `NAME`, and `REPS`. The `ENTRY` column must have a unique entry integer number for each treatment/genotype. The column `NAME` must have a unique name that identifies each treatment/genotype. The `REPS` column must have an integer entry for the replications of the checks and other entries. Both ENTRY and NAME must be unique, duplicates are not allowed. In the following, we show an example of the entries list format. This example has an entry list with four checks and 8 treatments/genotypes. It is crucial to allocate the checks in the top part of the file.

```{r, include=FALSE}
library(FielDHub)
ENTRY <- 1:12
NAME <- c(c("CHECK1","CHECK2","CHECK3"), paste0("G-", 4:12))
REPS <- c(rep(10,3),rep(1,9))
df <- data.frame(ENTRY,NAME,REPS)
```

```{r, echo = FALSE, results='asis'}
library(knitr)
kable(df)
```

1. Enter the number of stacked experiments in the **Input # of Stacked Expts** box. 

2. On the augmented RCBD we have the option to choose whether we randomize the entries or not, with the **Randomize Entries** toggle button. 

3. Enter the number of entries/treatments in the **Input # of Entries** box, which is 300 in our example experiment. 

4. Set the number of checks per block with the **Checks per Block** box

5. Set the number of blocks with **Input # of Blocks**.

The total number of plots in our field will be **Input # of Stacked Expts**(**Input # of Entries** + **Input # of Blocks** * **Checks per Block**), per location. 

6. Enter the number of locations in **Input # of Locations**. 

7. Select `serpentine` or `cartesian` in the **Plot Order Layout**. For this example we will set the `serpentine` layout.

8. To ensure that randomizations are consistent across sessions, we can set a seed number in the box labeled **Seed Number**. In this example, we will set it to 123. 

9. Enter the starting plot number in the **Starting Plot Number** box. If the experiment has multiple locations, you must enter a comma separated list of numbers the length of the number of locations for the input to be valid. 

10. Once we have entered the information for our experiment on the left side panel, click the **Run!** button to run the design. 

11. You will then be prompted to select the dimensions of the field from the list of options in the drop down in the middle of the screen with the box labeled **Select dimensions of field**. In our case, we will select 29 x 15. 

12. Click the **Randomize!** button to randomize the experiment with the set field dimensions and to see the output plots. If you change the dimensions again, you must re-randomize.

If you change any of the inputs on the left side panel after running an experiment initially, you have to click the Run and Randomize buttons again, to re-run with the new inputs.

## Outputs
After you run a single diagonal arrangement in FielDHub and set the dimensions of the field, there are several ways to display the information contained in the field book. The first tab, **Get Random**, shows the option to change the dimensions of the field and re-randomize, as well as a reference guide for experiment design. 

### Input Data
On the second tab, **Input Data**, you can see all the entries in the randomization in a list, as well as a table of the checks with the number of times they appear in the field. In the list of entries, the reps for each check is included as well.


### Randomized Field

The **Randomized Field** tab displays a graphical representation of the randomization of the entries in a field of the specified dimensions. The checks are all colored uniquely, showing the number of times they are distributed throughout the field. The display includes numbered labels for the rows and columns. You can copy the field as a table or save it directly as an Excel file with the _Copy_ and _Excel_ buttons at the top. 

### Plot Number Field
On the **Plot Number Field** tab, there is a table display of the field with the plots numbered according to the Plot Order Layout specified, either _serpentine_ or _cartesian_. You can see the corresponding entries for each plot number in the field book. Like the Randomized Field tab, you can copy the table or save it as an Excel file with the _Copy_ and _Excel_ buttons. 

### Field Book
The **Field Book** displays all the information on the experimental design in a table format. It contains the specific plot number and the row and column address of each entry, as well as the corresponding treatment on that plot. This table is searchable, and we can filter the data in relevant columns. 

# Function : `RCBD_augmented()`

You can run the same design with a function in the FielDHub package, `RCBD_augmented()`. We can enter the information describing the above design like this:

```{r, echo = TRUE, warning=FALSE, comment=''}
rcbd_a <- RCBD_augmented(
  lines = 35,
  checks = 3,
  b = 5, 
  repsExpt = 3, 
  nrows = 5, #number of rows in the field
  ncols = 10, #number of columns in the field
  seed = 123,
  random = TRUE,
  exptName = "Expt1"
)
```


```{r, comment='',echo=FALSE}
print(rcbd_a)
```

This returns a list consisting of all the information displayed in the output tabs in the FielDHub app: design information, plot layout, plot numbering, entries list, and field book. These are accessible by the `$` operator, i.e. `rcbd_a$layoutRandom` or `rcbd_a$fieldBook`. 

`rcbd_a$fieldBook` is a list containing information about every plot in the field, with information about the location of the plot and the treatment in each plot. As seen in the output below, the field book has columns for `ID`, `EXPT`, `LOCATION`, `YEAR`, `PLOT`, `ROW`, `COLUMN`, `CHECKS`, `ENTRY`, and `TREATMENT`.
```{r, comment='', echo=FALSE}
print(head(rcbd_a$fieldBook))
```


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
