---
title: "Scientific validation and computational limits"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Scientific validation and computational limits}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(FielDHub)
```

## What the checks establish

FielDHub uses three complementary checks: complete fixed-seed snapshots,
independent structural invariants, and independently calculated numerical
references. Snapshots detect changes; they do not establish scientific validity.
The invariant tests use explicit expected counts from the input fixtures, not
counts inferred from the generated design's summaries. Mutation tests check
that the independent helpers reject missing, repeated, or relabeled units.

The 40-case catalogue covers every design engine, including generated and
supplied labels, multiple locations, unequal replication, checks, and fillers.
`tests/testthat/test_scientific_invariants.R` checks the following properties on
every test platform; it is not restricted to the snapshot reference platform.

| Design family | Independently checked property |
|---|---|
| Completely randomized | Requested treatment multiplicities, including unequal replication |
| Complete blocks | Every test entry once per replicate and location; requested extra copies of checks |
| Latin squares | Every treatment once in each row and column of every square |
| Full factorial | Every requested factor combination in every replicate and location |
| Split and split-split plots | Complete factor combinations and a single whole-plot treatment within each physical whole plot |
| Strip plots | Complete horizontal-by-vertical combinations per replicate and location |
| Incomplete blocks and lattices | Complete treatment replicates, fixed block sizes, distinct nested units, and block efficiency |
| Row-column | Complete treatment replicates and complete replicate grids; separate SVD-based joint-efficiency tests |
| Diagonal and optimized arrangements | Experimental-entry multiplicities, total check allocation, and filler counts |
| Augmented complete blocks | Unreplicated test entries and every check in every block |
| Partially replicated | Requested one- and two-copy entry groups, excluding fillers |
| Sparse and multi-location allocations | Across-location copy budgets in balanced fixtures, site-level restrictions, and agreement between count matrices and entry lists |
| Family splitting | Every original entry retained; family counts across locations differ by at most one |
| Pair swapping | Entry frequencies and geometry retained; reported distances agree with `stats::dist()` |

All 36 field-design catalogue cases also have distinct final
`LOCATION`/`ROW`/`COLUMN` coordinates. `PLOT` is not a universal unique key:
split-unit designs can share a whole-plot number, fillers can have `PLOT = 0`,
and identifiers can restart at each location. Use the experimental hierarchy
and final physical coordinates appropriate to the design.

For example, a complete-block and physical-coordinate check can be written
without calling a package validator:

```{r complete-block-check}
design <- RCBD(t = 6, reps = 3, seed = 42)
book <- design$fieldBook
stopifnot(all(table(book$TREATMENT, book$REP) == 1L))
layout <- field_layout(design)
stopifnot(!anyDuplicated(layout[c("LOCATION", "ROW", "COLUMN")]))
stopifnot(identical(reproduce_design(design), design))
```

The complete-block, Latin-square, and full-factorial properties follow the
definitions in the NIST/SEMATECH handbook:
[randomized blocks](https://www.itl.nist.gov/div898/handbook/pri/section3/pri332.htm),
[Latin squares](https://www.itl.nist.gov/div898/handbook/pri/section3/pri3321.htm), and
[full factorials](https://www.itl.nist.gov/div898/handbook/pri/section3/pri333.htm).

## Numerical reference checks

For equally replicated incomplete-block designs, the independent test oracle
forms the treatment-by-block incidence matrix $N$ and computes

$$C = \operatorname{diag}(r) - N\operatorname{diag}(1/k)N^\mathsf{T}.$$

Here $r$ contains treatment replications and $k$ contains block sizes. If the
design is connected, the harmonic mean of its nonzero information eigenvalues,
divided by the common replication, gives the relative A-efficiency. The oracle
is checked against complete blocks (efficiency 1), the six pairs of four
treatments (efficiency 2/3), and a disconnected design (efficiency 0).
It then checks the reported efficiencies of incomplete-block, alpha, square,
and rectangular lattice catalogue cases. The legacy `blocksModel` table for
these functions describes the first location. The independent row-column
oracle instead uses an SVD projection to handle rank-deficient and crossed
row/column effects; see `test_row_column.R`.

The underlying optimization package documents its
[block models and efficiency reports](https://search.r-project.org/CRAN/refmans/blocksdesign/html/blocks.html).
These checks validate achieved designs, not a claim of global optimality or
uniform sampling over all possible randomizations.

Spatial simulation tests compare the separable AR1 construction against an
independent dense covariance/Cholesky calculation, including negative
correlations and nonsquare fields (`test_separable_spatial.R`). Response tests
check the stated treatment-specific model and unequal replication
(`test_truncated_response_contract.R` and `test_simulation.R`). These are
computational checks, not evidence that simulated responses fit a particular
biological experiment.

## Search bounds and diagnostics

| Operation | Bound and behavior at the limit |
|---|---|
| Latin-square placement | At most 100,000 placement iterations per square; a `fieldhub_search_error` records the square and budget on exhaustion |
| Repeated-check placement in RCBD | At most 100 attempts per block; warns and falls back to unrestricted placement if stratification cannot be completed |
| Pair-swap optimization | A finite geometry-dependent threshold range, at most `stop_iter` sweeps per threshold, and bounded sampled candidates per swap; diagnostics distinguish exhaustion, completion, and an empty range |
| Row-column search | Positive finite whole-number `iterations`; defaults are 200 external onestage searches or 1,000 twostage row-swap iterations |
| Allocation balancing | At most one upward pass through the allocation rows; warns if no eligible entry can complete the balancing |
| Dimension cache | At most 64 retained queries and 4 MiB of retained keys and results; temporary search memory is outside this bound |

Finite iteration counts are not wall-clock or memory guarantees. Work still
grows with field size, and external optimizer search internals are controlled
by `blocksdesign`. Forced allocation balancing can add entry copies to fill
smaller locations; inspect the returned allocation rather than assuming every
input copy budget is unchanged under that option. Pair-swap diagnostics
describe the retained result separately from an unsuccessful final attempt.

## Reproducibility and release evidence

Golden outputs are recorded on macOS. Structural and numerical tests run on
all supported test platforms, since optimizer choices can depend on R and
dependency versions and floating-point behavior. Replay comparisons use the
same package versions and recorded RNG settings. Review and explain any
deliberate fixed-seed change before accepting updated snapshots.

The release benchmark scripts under `tools/` measure optimizers, spatial
simulation, dimension searches, and layout exports. They record package/R
versions, platform, fixed inputs, elapsed time, and available allocation
measurements. Largest observed R allocation is not peak resident memory.
Compare results on the same host with comparable load, investigate changes,
and follow `RELEASING.md`; a successful local check does not establish that the
remote CI matrix or a deployment has passed.
