---
title: "Single Diagonal Arrangement"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{single_diagonal}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# App
One of the kinds of unreplicated designs you can use in FielDHub is single diagonal arrangement. In this design you can set the number of entries and the number of checks for the experiment. You can also choose to run the same experiment over multiple locations. For this experiment, we have 400 entries in the field and we want to use 6 checks. We will run this experiment in 1 location. 

## Inputs
First, make sure **Import entries' list** has "No" selected, as we are entering our own information about the experiment. 
Then, we can enter the number of entries in the **Input # of Entries** box, which is 400 in our case. Then, select 6 from the drop down on the **Input # of Checks** box. Since we want to run this experiment over 1 location, set **Input # of Locations** to 1. 

To ensure that randomizations are consistent across sessions, we can set a seed number in the box labeled **Seed Number**. For instance, we will set it to 123. 

Once we have entered the information for our experiment on the left side panel, click the **Run!** button to run the design. 

You will then be prompted to select the dimensions of the field from the list of options in the drop down in the middle of the screen with the box labeled **Select dimensions of field**. In our case, we will select 11 x 40. Click the **Randomize!** button to randomize the experiment with the set field dimensions and to see the output plots. 

If you change any of the inputs on the left side panel after running an experiment initially, you have to click the Run and Randomize buttons again, to re-run with the new inputs.

## Outputs
After you run a single diagonal arrangement in FielDHub and set the dimensions of the field, there are several ways to display the information contained in the field book. The first tab, **Expt Design Info**, shows the option to change the dimensions of the field and re-randomize, as well as a reference guide for experiment design. 

### Input Data
On the second tab, **Input Data**, you can see all the entries in the randomization in a list, as well as a table of the checks with the number of times they appear in the field. In our design, we get the following table of checks: 
```{r, include=FALSE}
ENTRY=1:6
NAME=c("CH1","CH2","CH3","CH4","CH5","CH6")
TIMES=c(7,7,6,7,6,7)
df=data.frame(ENTRY,NAME,TIMES)
library(knitr)
library(FielDHub)
```
```{r, echo = FALSE, results='asis'}
kable(df)
```


### Randomized Field
The **Randomized Field** tab displays a graphical representation of the randomization of the entries in a field of the specified dimensions. The checks are all colored uniquely, showing the number of times they are distributed throughout the field. The display includes numbered labels for the rows and columns. You can copy the field as a table or save it directly as an Excel file with the _Copy_ and _Excel_ buttons at the top. 

### Plot Number Field
On the **Plot Number Field** tab, there is a table display of the field with the plots numbered according to the Plot Order Layout specified, either _serpentine_ or _cartesian_. You can see the corresponding entries for each plot number in the field book. Like the Randomized field tab, you can copy the table or save it as an Excel file with the _Copy_ and _Excel_ buttons. 

### Field Book
The **Field Book** displays all the information on the experimental design in a table format. It contains the specific plot number and the row and column address of each entry, as well as the corresponding treatment on that plot. This table is searchable, and we can filter the data in relevant columns. 

# Function : `diagonal_arrangement()`

You can run the same design with a function in the FielDHub package, `diagonal_arrangement()`. We can enter the information describing the above design like this:

```{r, results='hide'}
diag <- diagonal_arrangement(
  nrows = 11,
  ncols = 40,
  lines = 400,
  checks = 6,
  l = 1,
  plotNumber = 1,
  seed = 123,
  exptName = "Expt1",
  locationNames = "FARGO"
)
```
```{r, echo = FALSE, warning=FALSE, comment='', include=FALSE}
diag <- diagonal_arrangement(
  nrows = 11,
  ncols = 40,
  lines = 400,
  checks = 6,
  l = 1,
  plotNumber = 1,
  seed = 123,
  exptName = "Expt1",
  locationNames = "FARGO"
)

```
```{r, comment='',echo=FALSE}
print(diag)
```
This returns a list consisting of all the information displayed in the output tabs in the FielDHub app: design information, plot layout, plot numbering, entries list, and field book. These are accessible by the `$` operator, i.e. `diag$layoutRandom` or `diag$fieldBook`. 

`diag$fieldBook` is a list containing information about every plot in the field, with information about the location of the plot and the treatment in each plot. As seen in the output below, the field book has columns for `ID`, `EXPT`, `LOCATION`, `YEAR`, `PLOT`, `ROW`, `COLUMN`, `CHECKS`, `ENTRY`, and `TREATMENT`.
```{r, comment='', echo=FALSE}
print(head(diag$fieldBook))
```


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
